---
title: "Process Mining Hip Replacement Pathways"
author: "Jenny Cooper"
date: "26/04/2021"
output: html_document

---
<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup, include=FALSE}


library(bupaR)
library(eventdataR)
library(xesreadR)
library(edeaR)
library(processmapR)
library(processmonitR)
library(processcheckR)
library(knitr)
library(kableExtra)
library(heuristicsmineR)
library(pm4py)
library(DiagrammeR)


#Other packages
library(tableone)
library(dplyr)
library(svgPanZoom)
library(DiagrammeRsvg)
library(processanimateR)
library(plyr)
library(ggplot2)
library(htmlwidgets)


#------------------------------------------------------------------------
hip_pathway <- readRDS( "hip_pathway_enriched.rds")
#------------------------------------------------------------------------
#Iterative cleaning steps for community data with similar service names
#------------------------------------------------------------------------

hip_pathway %>% 
   mutate(activity = replace(activity, 
                            activity=="community_district nursing service; long term conditions case management service; musculoskeletal service", "community_musculoskeletal service")) %>% 
  mutate(activity = replace(activity, 
                            activity=="community_district nursing service; podiatry service", "community_podiatry service")) -> hip_pathway
#------------------------------------------------------------------------
#Setting as eventlog format
#------------------------------------------------------------------------
hip_pathway %>% #a data.frame with the information in the table above
  eventlog(
    case_id = "Pseudo_NHS_Number",
    activity_id = "activity",
    activity_instance_id = "activity_instance",
    lifecycle_id = "status",
    timestamp = "timestamp",
    resource_id = "resource"
  ) -> hip_log
#------------------------------------------------------------------------
#Selected activities for the eventlog
#------------------------------------------------------------------------
selected_activity<- c("APPT_made",
                       "community_musculoskeletal service",
                       "community_physiotherapy service",
                       "elec_hip_replace",
                       "GP_refers_for_hip",
                       "OP_anaesthetic service",
                       "OP_diagnostic imaging service",
                       "OP_pain management service",
                       "OP_physiotherapy service",
                       "OP_trauma and orthopaedic service",
                       "primary_care_analgesic",
                       "primarycare_msk",
                       "OP_rheumatology service",
                       "AE")

restricted_activity<- c("APPT_made",
                      "community_musculoskeletal service",
                      "elec_hip_replace",
                      "GP_refers_for_hip",
                      "OP_diagnostic imaging service",
                      "OP_physiotherapy service",
                      "OP_trauma and orthopaedic service",
                      "primary_care_analgesic",
                      "primarycare_msk")

#Reduced from heuristics 0.5 model
selected_activity2<- c("APPT_made",
                       "community_musculoskeletal service",
                       "elec_hip_replace",
                       "GP_refers_for_hip",
                       "OP_anaesthetic service",
                       "OP_diagnostic imaging service",
                       "OP_pain management service",
                       "OP_physiotherapy service",
                       "OP_trauma and orthopaedic service",
                       "primary_care_analgesic",
                       "primarycare_msk",
                       "OP_rheumatology service",
                       "AE")

#------------------------------------------------------------------------
#Data cleaning for the cost variable
#------------------------------------------------------------------------
hip_log$cost2<- hip_log$cost
hip_log$cost2[is.na(hip_log$cost2)]<-0
#------------------------------------------------------------------------
#Mining algorithm
#------------------------------------------------------------------------
#Filtering event log for the selected activities
hip_log %>% filter_activity(selected_activity) -> selected_activity_log
#Ensuring additional levels are not in the model after reducing the activities (can become a problem for inductive miner)
selected_activity_log$activity<- factor(selected_activity_log$activity) 
selected_activity_log$status<- factor(selected_activity_log$status) 
#------------------------------------------------------------------------
#Filtering event log for the selected activities
#Reduced activities identified from the heuristics model (with 0.5 threshold)
hip_log %>% filter_activity(selected_activity2) -> selected_activity_log2 
#Ensuring additional levels are not in the model after reducing the activities (can become a problem for inductive miner)
selected_activity_log2$activity<- factor(selected_activity_log2$activity) 
selected_activity_log2$status<- factor(selected_activity_log2$status) 
#------------------------------------------------------------------------
#Inductive Miner log cleaning
#------------------------------------------------------------------------
selected_activity_log %>% 
  filter(status=="complete") -> patients_completes


patients_completes$activity<- factor(patients_completes$activity) 
patients_completes$status<- factor(patients_completes$status) 

#------------------------------------------------------------------------
```

### 1 Process Mining for Patients on a Hip Replacement Pathway
##### 1.1 Activity Frequency

The table below summarises the frequency and relative frequency of the activities involved in the hip replacement pathway identified from the event log.





```{r summary, echo = FALSE, fig.width=10, fig.height=5}
hip_log %>% filter_activity(selected_activity) %>% activity_frequency(level = "activity") %>% kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))

hip_log %>% filter_activity(selected_activity) %>% activity_frequency(level = "activity") %>% plot()


```





##### 1.2 Process Maps


The identified activities in the event log can be plotted in a process map which can be annotated with different information. For instance the process map below shows the frequencies of activities and flows. This shows a restricted set of activities for illustration purposes.





```{r frequency, out.width="100%", out.height="100%", echo = FALSE}
hip_log %>% 
  filter_activity(restricted_activity) %>% 
  process_map() %>%   
  DiagrammeRsvg::export_svg() %>% 
  svgPanZoom::svgPanZoom()
```





The next process map shows the event log from a performance perspective i.e. summarising the processing time of activities taking the median in this instance). This shows a restricted set of activities for illustration purposes.






```{r Performance, out.width="100%", out.height="100%", echo = FALSE}
hip_log %>% 
  filter_activity(restricted_activity) %>% 
  process_map(type=performance(median, units = "days")) %>% 
  DiagrammeRsvg::export_svg() %>% 
  svgPanZoom::svgPanZoom()

```





User defined measures can also be included. The process map below shows the average cost (median) of the different activities as well as the frequencies between them. This shows a restricted set of activities for illustration purposes.





```{r cost, out.width="100%", out.height="100%", echo = FALSE}
hip_log %>% 
  filter_activity(restricted_activity) %>% 
  process_map(type_nodes=custom(attribute="cost2", median, units = "£")) %>% 
  DiagrammeRsvg::export_svg() %>% 
  svgPanZoom::svgPanZoom()

```





The final process map shows all the selected activities and combines both frequencies and processing time.





```{r mix, out.width="100%", out.height="100%", echo = FALSE}
hip_log %>% 
  filter_activity(selected_activity) %>%
  process_map(type_nodes = frequency("relative_case"),
              type_edges = performance(median, units = "days")) %>% 
  DiagrammeRsvg::export_svg() %>% 
  svgPanZoom::svgPanZoom()
  

```





##### 1.3 Animated Process Maps

Each token in the process map below represents a patient following their own trace (or pathway) over time.
The tokens can be coloured by different attributes for instance here orange reflects patients who are 50 and over and yellow tokens represent patients less than 50.






```{r animated process map, echo = FALSE, warning = FALSE}

#By Age Group
hip_log %>% filter_activity(restricted_activity) %>%
  animate_process(mode='relative',
                  duration = 120,
                  jitter=10, 
                  legend = "color",
                  mapping = token_aes(
                    color= token_scale("age_group"
                                       , scale = "ordinal"
                                       , range = RColorBrewer::brewer.pal(2, "Spectral"))),
                  sizingPolicy = htmlwidgets::sizingPolicy(browser.fill = TRUE, viewer.fill = TRUE, browser.padding = 0,
    knitr.figure = FALSE, knitr.defaultWidth = 1600, knitr.defaultHeight = 1000) )

```





##### 1.4 Process Discovery Heuristics Miner


The plot below shows the dependency values for pairs of activities. These are used for setting the dependency thresholds for the model.






```{r dependency thresholds, echo = FALSE, fig.width=12, fig.height=8, warning = FALSE}

plot(dependency_matrix(selected_activity_log, threshold = 0.0, 
                       dependency_type = dependency_type_fhm(all_connected = TRUE)))


```





The following creates a dependency graph/matrix for the Heuristics Miner algorithm using a dependency threshold of 0.5.






```{r heuristics miner threshold 0.5, out.width="100%", out.height="100%",  echo = FALSE, warning = FALSE}

dependency_matrix(selected_activity_log, threshold = 0.5) %>% 
  render_dependency_matrix()%>% 
  DiagrammeRsvg::export_svg() %>% 
  svgPanZoom::svgPanZoom()


```





The following creates a dependency graph/matrix for the Heuristics Miner algorithm but using a dependency threshold of 0.6. We see that not all the activities are joined up hence the 0.5 threshold is taken forward.





```{r heuristics miner threshold 0.6, out.width="100%", out.height="100%", echo = FALSE, warning = FALSE}

dependency_matrix(selected_activity_log, threshold = 0.6) %>% 
  render_dependency_matrix()%>% 
  DiagrammeRsvg::export_svg() %>% 
  svgPanZoom::svgPanZoom()


```






There are other parameters to the Heuristics Miner algorithm including from the Flexible Heuristics Miner (FHM) 'all_connected' argument. This means the best antecedent and consequent (as determined by the dependency measure) will be added regardless of the threshold value [See heuristicsmineR documentation](https://cran.r-project.org/web/packages/heuristicsmineR/index.html)






```{r heuristics miner threshold all connected, out.width="100%", out.height="100%", echo = FALSE, warning = FALSE}

dependency_matrix(selected_activity_log, threshold = 0.6, 
                  dependency_type = dependency_type_fhm(all_connected = TRUE)) %>% 
  render_dependency_matrix()%>% 
  DiagrammeRsvg::export_svg() %>% 
  svgPanZoom::svgPanZoom()


```





The plot below shows the causal graph otherwise known as the heuristics net for a dependency threshold of 0.5.





```{r heuristics net 0.5, out.width="100%", out.height="100%", echo = FALSE, warning = FALSE}


causal_net(selected_activity_log, threshold = .5) %>% render_causal_net() %>% 
  DiagrammeRsvg::export_svg() %>% 
  svgPanZoom::svgPanZoom()


```






##### 1.5 Process Discovery Inductive Miner

The plots below show the resulting petri net discovered using the inductive mining algorithm (inductive miner directly follows algorithm) and after rendering to DOT format.

The process model discovered using the inductive mining algorithm consists of a petri net, an initial marking and a final marking.






```{r inductive miner, echo = FALSE, out.width="50%", out.height="40%", warning = FALSE}

pn <- discovery_inductive(patients_completes)

# This results in an auto-converted bupaR Petri net and markings
str(pn)
class(pn$petrinet)

# Render with bupaR
render_PN(pn$petrinet) %>%DiagrammeRsvg::export_svg() %>% 
  svgPanZoom::svgPanZoom()

```







```{r inductive miner 2, echo = FALSE, out.width="50%", out.height="50%", warning = FALSE}

# Render with  PM4PY and DiagrammeR
viz <- reticulate::import("pm4py.visualization.petrinet")

# Convert back to Python
py_pn <- r_to_py(pn$petrinet)
class(py_pn)

# Render to DOT with PMP4Y
dot <- viz$factory$apply(py_pn)$source
grViz(diagram = dot) %>%
  DiagrammeRsvg::export_svg() %>% 
  svgPanZoom::svgPanZoom()
  


```







We can use the event log data to compute the alignments by comparing with the model outputs. The alignment table can be used to compare the log_label and the model_label to determine when a move is synchronous.

The quality of the model can be determined by replaying the eventlog data over the discovered model.
Fitness in this instance is 100% which means the model fits the data well. 

Other measures are also calculated including precision, generalisation, simplicity, average weight and fscore.
[pm4py documentation for quality measures](https://pm4py.fit.fraunhofer.de/documentation#evaluation "pm4py documentation")







```{r inductive miner 3, echo = FALSE, warning = FALSE}

# Render with  PM4PY and DiagrammeR
quality <- evaluation_all(patients_completes, pn$petrinet, pn$initial_marking, pn$final_marking)

print(quality)

```





